---
title: "Coursera: Matrix Algebra for Engineers. Week 1"
output: html_notebook
---

R notebook understands latex math symbols.

$$a_1b_2 - a_2b_1$$


```{r}
A <- matrix(c("a","b","c","d"), nrow = 2, byrow = "FALSE")

```
The main diagonal of a matrix A are the entries 
$a_{ij}$ where $i = j$

(a) Write down the three-by-three matrix with ones on the diagonal and zeros elsewhere.

```{r}
matrix(c(1,0,0,
         0,1,0,
         0,0,1), nrow = 3)
```

(b) Write down the three-by-four matrix with ones on the diagonal and zeros elsewhere.
```{r}
matrix(c(1,0,0,0,
         0,1,0,0,
         0,0,1,0), nrow = 3, byrow = TRUE)
```

(c) Write down the four-by-three matrix with ones on the diagonal and zeros elsewhere.
```{r}
matrix(c(1,0,0,
         0,1,0,
         0,0,1,
         0,0,0), nrow = 4, byrow = TRUE)
```

Define the matrices

```{r}
A <- matrix(c(2,1,1,-1,-1,1), nrow = 2)
B <- matrix(c(4,2,-2,-4,1,-2), nrow = 2)
C <- matrix(c(1,2,2,1), nrow = 2)
D <- matrix(c(3,4,4,3), nrow = 2)
E <- matrix(c(1,2), nrow = 2)
```
compute if defined

```{r}
B-2*A
```

```{r}
3*C-E
```

```{r}
A%*%C
```


```{r}
C%*%D
```


```{r}
C%*%B
```

```{r}
A <- matrix(c(1,2,2,4), nrow = 2)
B <- matrix(c(2,1,1,3), nrow = 2)
C <- matrix(c(4,0,3,2), nrow = 2)

```

```{r}
A%*%B
```


```{r}
A%*%C
```


```{r}
B == C
```


```{r}
identical(B,C)
```


```{r}
all.equal(B,C)
```




```{r}
A <- matrix(c(1,1,1,1,2,3,1,3,4), nrow = 3)
D <- matrix(c(2,0,0,0,3,0,0,0,4), nrow = 3)
```

```{r}
A%*%D
```


```{r}
D%*%A
```

Prove the associative law for matrix multiplication. 
That is, let 
A be an m-by-n matrix, 
B an n-by-p matrix, and 
C a p-by-q matrix.

Prove that `A(BC) = (AB)C`

```{r}
m <- 3
n <- 4
p <- 5
q <- 6
A <- matrix(rnorm(m*n), nrow = m)
B <- matrix(rnorm(n*p), nrow = n)
C <- matrix(rnorm(p*q), nrow = p)

```

```{r}
A %*% (B %*% C) 
```

```{r}
(A %*% B) %*% C
```

